---
title: "ICA Tuesday"
author: "Aditya Jain"
date: "2022-09-14"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1 

### a.

Modify and run the board game example simulation (without bonus) so that it runs 100,000 times. Give the percentage each of the 3 players wins. All problems below will be run with 100,000 simulations.

```{r }
sim <- 100000
win <- replicate(sim,0)
for(i in 1:sim){
  P <- c(0,0,0)
  while(max(P) < 8){
    for(j in 1:3){
      roll <- sample(1:6,1)
      P[j] = P[j] + roll
      if(P[j] >= 8){
        break
      }
    }
  }
  win[i] <- j
}
print(paste("Probability that player one wins",round(100*(sum(win == 1)/sim),1),"%"))
print(paste("Probability that player two wins",round(100*(sum(win == 2)/sim),1),"%"))
print(paste("Probability that player third wins",round(100*(sum(win == 3)/sim),1),"%"))

```
\newpage

### b.

Modify and run the board game example simulation so that the winner is the first to 50. How do the winning percentages compare to (a)? Why?

```{r}
sim <- 100000
win <- replicate(sim,0)
for(i in 1:sim){
  P <- c(0,0,0)
  while(max(P) < 50){
    for(j in 1:3){
      roll <- sample(1:6,1)
      P[j] = P[j] + roll
      if(P[j] >= 50){
        break
      }
    }
  }
  win[i] <- j
}

print(paste("Probability that player one wins",round(100*(sum(win == 1)/sim),1),"%"))
print(paste("Probability that player two wins",round(100*(sum(win == 2)/sim),1),"%"))
print(paste("Probability that player third wins",round(100*(sum(win == 3)/sim),1),"%"))
```
The percentage moves closer to fair distribution, where winning chances are equal for all (33%). This is becuase the game is played for longer duration as winning target is 50. This gives enough tries. 

\newpage

### c.

Modify and run the board game example simulation in (b) (winner is first to 50) so that a 4-sided die is used instead. How do the winning percentages compare to (b)? Why?

```{r}
sim <- 100000
win <- replicate(sim,0)
for(i in 1:sim){
  P <- c(0,0,0)
  while(max(P) < 50){
    for(j in 1:3){
      roll <- sample(1:4,1)
      P[j] = P[j] + roll
      if(P[j] >= 50){
        break
      }
    }
  }
  win[i] <- j
}

print(paste("Probability that player one wins",round(100*(sum(win == 1)/sim),1),"%"))
print(paste("Probability that player two wins",round(100*(sum(win == 2)/sim),1),"%"))
print(paste("Probability that player third wins",round(100*(sum(win == 3)/sim),1),"%"))
```
The winning percentage remains same as it is independent to the number of sides on the dice, in cases when larger amount of simulations are run. 

\newpage

### d 

Modify and run the board game example simulation in (a) so that if any player lands on the 4th square, they get an extra turn (play to 8).
```{r}
sim <- 100000
win <- replicate(sim,0)
for(i in 1:sim){
  P <- c(0,0,0)
  while(max(P) < 8){
    for(j in 1:3){
      roll <- sample(1:6,1)
      P[j] = P[j] + roll
      if(P[j] == 4){
        roll <- sample(1:6,1)
        P[j] = P[j] + roll
      }
      if(P[j] >= 8){
        break
      }
    }
  }
  win[i] <- j
}

print(paste("Probability that player one wins",round(100*(sum(win == 1)/sim),1),"%"))
print(paste("Probability that player two wins",round(100*(sum(win == 2)/sim),1),"%"))
print(paste("Probability that player third wins",round(100*(sum(win == 3)/sim),1),"%"))
```
\newpage

### e.

Modify and run the board game example simulation in (d) so that there are 5 players. Plot a histogram of the wins by each player.

```{r}
sim <- 100000
win <- replicate(sim,0)
for(i in 1:sim){
  P <- c(0,0,0,0,0)
  while(max(P) < 8){
    for(j in 1:5){
      roll <- sample(1:6,1)
      P[j] = P[j] + roll
      if(P[j] == 4){
        roll <- sample(1:6,1)
        P[j] = P[j] + roll
      }
      if(P[j] >= 8){
        break
      }
    }
  }
  win[i] <- j
}

hist(win)

```


