---
title: "STT - Mini Project"
author: "Aditya Jain"
header-includes:
   - \usepackage{bbm}
   - \usepackage{amsmath}
   - \usepackage[utf8]{inputenc}
   - \usepackage[english]{babel}
   - \usepackage{amsthm}
   - \DeclareUnicodeCharacter{2212}{\textendash}
date: "`r Sys.Date()`"
output:
  pdf_document:
    latex_engine: pdflatex
    toc : true
  html_document:
    df_print: paged
subtitle: Code Output File
---


```{r setup, include=FALSE}
library(knitr)
library(tidyverse)
library(Hmisc)
library(ggplot2)
library(reshape2)
library(gridExtra)
library(tidyverse, warn.conflict=F)
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, error = FALSE, warning = FALSE)
```

\newpage

## Question 1

**(a)**

The affinity matrix was created as shown below for 15 nodes. The below matrix has been created for just 1 simulation but each and every simulation would have it's own affinity matrix. The affinity matrix is of 15 columns and 15 rows which will give the information of which node is connected to which node. For ex- if on index (1,2) and (2,1) there is a "1" in affinity matrix, then that means there is a connection between nodes 1 and 2.

```{r, echo=FALSE}
#For 1 simulation
sim <- 1
nodes <- 15
maxedge <- replicate(sim,0)
maxloc <- replicate(sim,0)
mat <- replicate(nodes*nodes,0)
dim(mat) <- c(nodes,nodes)
for(i in 1:sim){
  edges <- replicate(nodes, 0)
  edges[1] = 1
  edges[2] = 1
  mat[1,2] <- 1
  mat[2,1] <- 1
  for(j in 3:nodes){
    sumedge <- sum(edges)
    edges[j] = 1
    new <- sample(1:(j-1),1, prob = edges[1:(j-1)]/sumedge)
    edges[new]=edges[new]+1
    mat[new,j] <- 1
    mat[j,new] <- 1
  }
  #print(mat)
}

```

```{r}
#For more than 1 simulations
sim <- 10000
nodes <- 15
mat <- replicate(sim*nodes*nodes,0)
dim(mat) <- c(nodes,nodes,sim)
for(i in 1:sim){
  edges <- replicate(nodes, 0)
  edges[1] <- 1
  edges[2] <- 1
  mat[1,2,i] <- 1
  mat[2,1,i] <- 1
  for(j in 3:nodes){
    sumedge <- sum(edges)
    edges[j] = 1
    new <- sample(1:(j-1),1, prob = edges[1:(j-1)]/sumedge)
    edges[new]=edges[new]+1
    mat[new,j,i] <- 1
    mat[j,new,i] <- 1
  }
}
```

**(b)**

The edge vector contains the information of how many edges are associated with that particular node. So, for ex- if edge[1] has the value of 6 then that means Node number 1 is associated with or connected with 6 edges/nodes. 

Therefore, for this question's sub-part 10,000 simulations was done and for each simulation the edge vector was stored, which was subsequently added to the previous simulation's edge vector to evaluate that how many edges has been connected with each node from Node 1-15 for total of 10,000 simulations. 

Observations-

The number of edges each node has was not equal. When we plotted the distribution plot i.e. histogram, between number of edges connected to it and node number, it was observed that the relationship is an exponential distribution curve. This is shown below.

It can be seen that maximum edges was connected to Initial Nodes i.e. Node 1,2,3 and 4. The total edges connected with node 1 and node 2 are approximately same and subsequently after that the number of edges decreases with the increase in the Node number. 

```{r,fig.width=7,fig.height=3}
sim <- 10000
nodes <- 15
tot_edges <- replicate(nodes,0)
for(i in 1:sim){
  edges <- replicate(nodes, 0)
  edges[1] = 1
  edges[2] = 1
  for(j in 3:nodes){
    sumedge <- sum(edges)
    edges[j] = 1
    new <- sample(1:(j-1),1, prob = edges[1:(j-1)]/sumedge)
    edges[new[1]]=edges[new]+1
  }
  tot_edges <- tot_edges+edges
}
Node_number<-seq(1,15)
df<- data.frame(tot_edges,Node_number)
p<-ggplot(data=df, aes(x=Node_number, y=tot_edges, fill=Node_number)) + geom_bar(stat="identity")+theme_minimal()+ labs(title = "Relationship between edge number and number of edges",x="Node Number", y="Number of edges")
p
```
**(c)**

```{r}
sim <- 10000
nodes <- 15
tot_edges <- replicate(nodes,0)
for(i in 1:sim){
  edges <- replicate(nodes, 0)
  edges[1] = 1
  edges[2] = 1
  for(j in 3:nodes){
    sumedge <- sum(edges)
    edges[j] = 1
    new <- sample(1:(j-1),1, prob = edges[1:(j-1)]/sumedge)
    edges[new[1]]=edges[new]+1
  }
  tot_edges <- tot_edges+edges
}
sum<-0
count<-0
for( i in 1:nodes){
  sum<-sum+tot_edges[i]*i
  count<-count+tot_edges[i]
}
mean<-sum/count
mean
count2<-0
q<-33/100*nodes
for( i in 1:q){
  count2<-count2+tot_edges[i]
}
count2/count*100
```

It can be seen that initial 1/3rd of the total nodes, i.e. for 15 nodes it is 4.95 (approx. 5), is associated with 50% of the total edges formed and the mean value is approximately 6.

```{r,fig.width=6,fig.height=2}
sim <- 10000
nodes <- 15
mat <- replicate(sim*nodes*nodes,0)
dim(mat) <- c(nodes,nodes,sim)
#tot_edges <- replicate(nodes,0)
for(i in 1:sim){
  edges <- replicate(nodes, 0)
  edges[1] <- 1
  edges[2] <- 1
  mat[1,2,i] <- 1
  mat[2,1,i] <- 1
  for(j in 3:nodes){
    sumedge <- sum(edges)
    edges[j] = 1
    new <- sample(1:(j-1),1, prob = edges[1:(j-1)]/sumedge)
    edges[new]=edges[new]+1
    mat[new,j,i] <- 1
    mat[j,new,i] <- 1
  }

  #tot_edges <- tot_edges+edges
}
nodes_sum<-rowSums(mat,dims=1)
m<-nodes_sum/(nodes*sim)
Node_number<-seq(1,15)
df2<- data.frame(m,Node_number)
p<-ggplot(data=df2, aes(x=Node_number, y=m, group=1)) + geom_line(colour="red")+ geom_point()
p+labs(title="Plot of mean of multiple simulations",x ="Node Number", y = "Mean")
```

For multiple simulations we evaluated the mean for each node and then plotted the distribution using the affinity matrix. So, for each node and for all simulations total number of connections were calculated and divided by the total count which finally gave the mean value for each node. It was found that the mean value is also distributed exponentially and there is a decrease in it's value as the number of nodes is increased.

**(d)**

```{r}
sim <- 10000
nodes <- 100
maxedge <- replicate(sim,0)
maxloc <- replicate(sim,0)
mat <- replicate(nodes*nodes,0)
dim(mat) <- c(nodes,nodes)
tot_edges <- replicate(nodes,0)
for(i in 1:sim){
  edges <- replicate(nodes, 0)
  edges[1] = 1
  edges[2] = 1
  mat[1,2] <- 1
  mat[2,1] <- 1
  for(j in 3:nodes){
    sumedge <- sum(edges)
    edges[j] = 1
    new <- sample(1:(j-1),1, prob = edges[1:(j-1)]/sumedge)
    edges[new[1]]=edges[new]+1
    mat[new,j] <- 1
    mat[j,new] <- 1
  }
  tot_edges <- tot_edges+edges
}
sum<-0
count<-0
for( i in 1:nodes){
  sum<-sum+tot_edges[i]*i
  count<-count+tot_edges[i]
}
mean<-sum/count
mean
count2<-0
q<-25/100*nodes
for( i in 1:q){
  count2<-count2+tot_edges[i]
}
count2/count*100
```

In this, it was found that approx. 1/4 of total nodes were associated with 50% of the total edges formed and the mean value is approximately 34.

In this sub-part of the question, the number of nodes were increased to 100 and again the same distribution graph was plotted. This plot made the picture clearer that the plot is an exponential distribution. This can be visualized below.

```{r,fig.width=6,fig.height=3}
Node_number<-seq(1,100)
df2<- data.frame(tot_edges,Node_number)
p<-ggplot(data=df2, aes(x=Node_number, y=tot_edges, fill=Node_number)) + geom_bar(stat="identity")+theme_minimal()+ labs(title = "Relationship between edge number and number of edges",x="Node Number", y="Number of edges")
p
```

```{r,fig.width=6,fig.height=2}
sim <- 1000
nodes <- 100
mat <- replicate(sim*nodes*nodes,0)
dim(mat) <- c(nodes,nodes,sim)
#tot_edges <- replicate(nodes,0)
for(i in 1:sim){
  edges <- replicate(nodes, 0)
  edges[1] <- 1
  edges[2] <- 1
  mat[1,2,i] <- 1
  mat[2,1,i] <- 1
  for(j in 3:nodes){
    sumedge <- sum(edges)
    edges[j] = 1
    new <- sample(1:(j-1),1, prob = edges[1:(j-1)]/sumedge)
    edges[new]=edges[new]+1
    mat[new,j,i] <- 1
    mat[j,new,i] <- 1
  }

  #tot_edges <- tot_edges+edges
}
nodes_sum<-rowSums(mat,dims=1)
m<-nodes_sum/(nodes*sim)
Node_number<-seq(1,100)
df2<- data.frame(m,Node_number)
p<-ggplot(data=df2, aes(x=Node_number, y=m, group=1)) + geom_line(colour="red")+ geom_point()
p+labs(title="Plot of mean of multiple simulations",x ="Node Number", y = "Mean")
```

It was found that the mean value is also distributed exponentially and there is a sharp decrease in it's value as the number of nodes is increased.

\newpage

## Question 2


```{r}
setwd('C:\\Users\\Adi\\Desktop\\Fall-22 Study Material\\STT 810\\Mini Proj -1')
nndb <- read.csv(file = "nndb_flat.csv",header =TRUE,  stringsAsFactors = TRUE,
                 encoding = 'UTF-8', na.strings = "" )
nndb <- as.data.frame(nndb)
```

Some basic information gathered from the internet about the data and the variables -

* The nndb data is a flattened version of the USDA National Nutrition Database
* Each record is a food composition data per 100 grams for all major source of food in USA

The columns are mostly self-explanatory. The nutrient columns end with the units, so:

*	Nutrient_g is in grams
*	Nutrient_mg is in milligrams
*	Nutrient_mcg is in micro grams
*	Nutrient_USRDA is in percentage of US Recommended Daily Allows (e.g. 0.50 is 50%)


```{r}
# Getting dimension and column names of the data
dimensions <- dim(nndb)
col_names <-  names(nndb)
paste(dimensions)
paste(col_names)
```

* Within the data set, their are 6 non numerical variables consisting of categorical and nominal data
* One Energy column, the total energy in kilo calorie per 100 gm of food
* Five basic nutrients like protein, fat, carb. etc in grams
* Vitamins and other minerals in mili and micro grams
* The rest of the columns contain values limited per day by US government 
* The data majorly contains numeric values from column energy_kcal to zinc_usrda 
* Nominal and categorical data are stored as Factor, with levels equal to number of unique data points
* Id is stored as int data type
* Common, Manufacturing and Scientific names are the only fields with missing or null values 


```{r, fig.height= 10, fig.width= 6}
# Correlation matrix for the numerical fields
cor_data_clean <- round(cor(nndb[,8:ncol(nndb)]),1)
data.frame(cor_data_clean) #%>% head() %>% kable()

```

* All the USRDA attributes are redundant as they are perfectly correlated, hence removing these features for further analysis

```{r, fig.height= 10, fig.width= 10}
# Correlation heat map for numerical fields
cor_data_clean <- cor(nndb[,8:30])

melted_cormat <- melt(data = cor_data_clean)

fig = ggplot(melted_cormat, aes(x = Var1, y = Var2, fill = value)) +
  geom_tile(color = "white"
            ) +
  xlab(" ") + ylab(" ")+
  labs(title = "Heatmap of Correlation values in NNDB Data Set")+
  scale_fill_gradient(low = "white", high = "blue") +
  geom_text(aes(label = round(value,2)), color = "Black", size = 2)+
  coord_fixed() +
  guides(fill = guide_colourbar(title = "Pearson Cor.",
                                barwidth = 0.5,
                                barheight = 20))+
  theme(axis.text.x = element_text(angle = 70, margin = margin(t = 25)))

fig

```

The correlation heat map indicates positive correlation between variables 


```{r,fig.height= 4, fig.width= 6}
# Understanding distribution of variables by creating box plots

proximates = subset(nndb, select=c("Protein_g","Fat_g","Carb_g",
                                  "Sugar_g","Fiber_g"))

melt_proximates = melt(proximates)

# Figure 1: Boxplots of proximates in all foods

ggplot(data=melt_proximates) +
  geom_boxplot(aes(x=variable, y=value)) +
  facet_wrap(~variable, scale="free", ncol=5) +
  labs(x="Proximates", y="Values")

```

The above box plots shows us the distribution of five main nutrients in food.

* We can clearly identify large numbers of outliers in every field, especially in sugar 
* All nutrient category contains few products which are way above the US daily accepted limit

```{r, fig.height= 4, fig.width= 6}

minerals = subset(nndb, select=c("Calcium_mg","Copper_mcg", "Iron_mg", "Magnesium_mg", "Phosphorus_mg", "Zinc_mg", "Manganese_mg", "Selenium_mcg"))

melt_minerals = melt(minerals)

# Figure 2: Boxplots of minerals in all foods

ggplot(data=melt_minerals) +
  geom_boxplot(aes(x=variable, y=value)) +
  facet_wrap(~variable, scale="free", ncol=4) +
  labs(x="Minerals", y="Values")
```

Minerals like phosphorous and selenium are present in huge amount in products, Way above accepted values by the US department.

```{r,fig.height= 3, fig.width= 6}
vitamin = subset(nndb, select=c("VitA_mcg","VitB6_mg","VitB12_mcg", "VitC_mg", "VitE_mg", "Folate_mcg","Niacin_mg","Riboflavin_mg","Thiamin_mg"))

melt_vitamin = melt(vitamin)


ggplot(data=melt_vitamin) +
  geom_boxplot(aes(x=variable, y=value)) +
  facet_wrap(~variable, scale="free", ncol=5) +
  labs(x="Vitamins", y="Values")
```


```{r, fig.height= 4, fig.width= 6}
# Scatter plot
ggplot(data = nndb, aes(x = Energy_kcal, y = Fat_g)) + 
  geom_point(size = 2)+ 
  labs(x = "Energy_Kcal", y = "Fat_g", title = "Energy vs Fat")
```

The scatter plot shows positive correlation between energy and fat. As visualized, as fat increases the energy Kcal also increases.

```{r,  fig.height= 4, fig.width= 6}
# PCA

pr.out = prcomp(nndb[,9:30], center=TRUE, scale=TRUE)  # With center=TRUE, scale=TRUE
pr.out$x = -pr.out$x
pr.out$rotation = -pr.out$rotation

pve = pr.out$sdev^2 / length(pr.out$sdev)  # Proportion of variance explained


# Figure 4: Proportion of variance explained and their cumulative sum from PCA

p1 = ggplot() +
  geom_line(aes(x=c(1:length(pr.out$sdev)), y=pve)) +
  geom_point(aes(x=c(1:length(pr.out$sdev)), y=pve), size=3) +
  geom_hline(yintercept=pve[7], color='red') +
  labs(x="Principal component", y="Proportion of variance explained")

p2 = ggplot() +
  geom_line(aes(x=c(1:length(pr.out$sdev)), y=cumsum(pve))) +
  geom_point(aes(x=c(1:length(pr.out$sdev)), y=cumsum(pve)), size=3) +
  geom_hline(yintercept=cumsum(pve)[7], color='red') +
  annotate("text", x=20, y=0.59, label=paste(round(cumsum(pve)[7],3))) +
  labs(x="Principal component", y="Cumulative proportion of variance explained")

grid.arrange(p1, p2, ncol=2)


```

Principal component analysis helps to extract features in our data set which explains high proportion of variance. By applying PCA on our data, we are able to explain 70 percent of variance with the help of only 7 components. 


```{r,  fig.height= 4, fig.width= 6}

###
### Selection of nutrients
###

pve.mat = matrix(rep(pve, each = 7), nrow=length(pve))
nutrient.impact = apply(pr.out$rotation[,c(1:7)]^2 * pve.mat, 1, sum)
melt_NI = melt(nutrient.impact)

# Figure 7: Contribution of nutrients on the first 7 prinicpal components

ggplot(data=melt_NI) +
  geom_col(aes(x=reorder(rownames(melt_NI), -value), y=value)) +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(x="Nutrients", y="Variable importance")

top6 = rownames(melt_NI)[order(-melt_NI$value)[1:6]]

```

The bar chart shows contribution of nutrients in the top 7 principal component.

```{r, fig.height= 4, fig.width= 6}
slopes = rep(NA, 6)

for(s in c(1:6)) {
  slopes[s] = pr.out$rotation[top6[s],2] / pr.out$rotation[top6[s],1]
}

# Figure 8: Biplot of 6 most significant nutrients impacting the first 6 principal components

ggplot() +
  geom_point(aes(x=pr.out$x[,1], y=pr.out$x[,2])) +
  geom_segment(aes(x=0, y=0, xend=-5, yend=slopes[1]*(-5)), 
               arrow = arrow(length = unit(0.03, "npc")), color="red") +
  annotate("text", x=-4, y=slopes[1]*(-5)+1, label=top6[1], color="red") +
  geom_segment(aes(x=0, y=0, xend=10, yend=slopes[2]*10), 
               arrow = arrow(length = unit(0.03, "npc")), color="red") +
  annotate("text", x=13, y=slopes[2]*10, label=top6[2], color="red") +
  geom_segment(aes(x=0, y=0, xend=12, yend=slopes[3]*12), 
               arrow = arrow(length = unit(0.03, "npc")), color="red") +
  annotate("text", x=12, y=slopes[3]*12-1, label=top6[3], color="red") +
  geom_segment(aes(x=0, y=0, xend=15, yend=slopes[4]*15), 
               arrow = arrow(length = unit(0.03, "npc")), color="red") +
  annotate("text", x=17, y=slopes[4]*15+1, label=top6[4], color="red") +
  geom_segment(aes(x=0, y=0, xend=3, yend=slopes[5]*3), 
               arrow = arrow(length = unit(0.03, "npc")), color="red") +
  annotate("text", x=3, y=slopes[5]*3-1, label=top6[5], color="red") +
  geom_segment(aes(x=0, y=0, xend=30, yend=slopes[6]*30), 
               arrow = arrow(length = unit(0.03, "npc")), color="red") +
  annotate("text", x=30, y=slopes[6]*30+2, label=top6[6], color="red") +
  labs(x="PC1", y="PC2")
```

The biplot shows 6 most significant nutrients impacting the first two principal components. 

* It is observed that vitamins, protein, carb and fibers are the best indicators of variance in the data